---
title: "Neotoma Workshop, AmQua July 2016"
author: "Jack Williams, Simon Goring, and Eric Grimm"
date: "July 2, 2016"
output: html_document
---

# Neotoma Workshop, AmQua July 2016
## Useful Websites
### Neotoma Resources
* SERC Carleton: <http://serc.carleton.edu/neotoma/activities.html>
* Neotoma APIs: <http://api.neotomadb.org/>
* Neotoma Manual and Database Structure: <http://neotoma-manual.readthedocs.io/en/latest/>
* Neotoma Entity Relationship Diagram: <http://www.neotomadb.org/uploads/NeotomaDMD.pdf>
* Goring et al. 2015 neotoma package: <http://www.openquaternary.com/articles/10.5334/oq.ab/>

### Guides: RMarkdown & GitHub
* RMarkdown Reference Guide: <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>
* Using GitHub with R and RStudio: <http://www.molecularecologist.com/2013/11/using-github-with-r-and-rstudio/>
* RStudio Support:  <https://support.rstudio.com/hc/en-us/articles/200532077?version=0.99.489&mode=desktop>

## Workshop Activities and Background
### Part 1: Explorer Demo
* Walkthrough of Neotoma Explorer.  Exercise available at SERC Carleton: <http://serc.carleton.edu/neotoma/activities.html>
* save JSON and drag back to Explorer

### Part 2:  APIs & JSON
#### Intro to APIs
Neotoma makes its data available through a set of programming tools known as APIs (Application Programming Interfaces). An API allows different applications to talk to one another. In most cases you will see an API being used in a web environment, where commands are passed through the URL using particular, fixed, variables.

Usually, if you want to analyze many sites or pass Neotoma data to other analytical environments, such as R, Matlab, or Python, you will want to use the APIs.  

Here, we'll introduce the concept of APIs and show how the Neotoma APIs work. In the next section, we will explore the _neotoma_ R package.  This package sits on top of the Neotoma APIs and uses them to get data and metadata from the Neotoma database and pull them into Neotoma.

Let's begin with an example of an API that we use all the time:  Google. Take a second right now to go to Google and type in a search term, then hit enter. You should see something like this:

> https://www.google.com/webhp?sourceid=chrome-instant&ion=1&espv=2&ie=UTF-8#q=when%20did%20the%20quaternary%20begin

This is the API that interfaces between your web browser and Google's databases.

You can see I searched for the string “when did the quatnerary begin”, it is prefaced by “q=”. Google's servers know that when they are reading the URL everything followed by the hash sign (#) is part of a set of variables.

#### Neotoma APIs
[Neotoma](http://www.neotomadb.org) has the same kind of API, although it is specialized to serve up the paleoecological records stored in the database. The APIs are documented and described on the Neotoma developer [website](http://api.neotomadb.org/)



Scrap text from standard RMarkdown template
```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
